---
title: "Stomach cancer paper Figure 2 ABC"
author: "Vi Dang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    css: font.css
---

```{r setup, include = FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r }
# Load libraries
library(Seurat)
library(tidyverse)
library(SingleCellExperiment)
library(scater)
library(patchwork)
library(RColorBrewer)
library(factoextra)
library(pheatmap)
library(ggplot2)
library(ggforce)
library(ggrepel)
library(ggvenn)
library(org.Hs.eg.db)
library(msigdbr)
library(fgsea)
library(clusterProfiler)
library(DESeq2)
```

```{r }
# Specify the path
main_path <- '/group/poetsch_projects/poetsch_sc/Stomach_paper_repository'
rds_path  <- file.path(main_path,'RDS')
txt_files <- file.path(main_path,'txt_files')
plot_path <- file.path(main_path,"Figures")
figures_data <- file.path(main_path,"Figures_data")
dir.create(figures_data)
```
# Load data and edit sample meta data 
```{r }
# Load seurat object containing data from all 9 samples integrated and clustered (for pseudobulk DE later)
seurat_integrated <- readRDS(file.path(rds_path,'clustered_0.35_integrated_seurat_ref.rds'))
seurat_integrated
```

```{r }
#Edit the meta data
seurat_integrated$patient<- str_extract(seurat_integrated$orig.ident,"^[^[_-]]*")
seurat_integrated$treatment<-str_extract(seurat_integrated$orig.ident,"[^-]*$")
meta <- seurat_integrated@meta.data
meta <- meta %>%
  mutate(sample = case_when(
    grepl("-no", orig.ident) ~ sub("-no", " Untreated", orig.ident),
    grepl("-vitro", orig.ident) ~ sub("-vitro", " Ex vivo", orig.ident),
    grepl("-vivo", orig.ident) ~ sub("-vivo", " In vivo", orig.ident),
    TRUE ~ orig.ident
  ), 
   Treatment = case_when(
    treatment == "no" ~ "Untreated",
    treatment == "vitro" ~ "Ex vivo",
    treatment == "vivo" ~ "In vivo"
   ))
seurat_integrated@meta.data <- meta
```

# Transcriptomic dissimilarity
## Aggregate read counts per sample and calculate transcriptomic dissimilarity
```{r }
#Transform to SCE for bulk analysis
DefaultAssay(seurat_integrated) <-'RNA'
counts_matrix <- GetAssayData(object = seurat_integrated, slot = "counts")
coldata <- seurat_integrated@meta.data
#umap_embeddings<-Embeddings(object = seurat_integrated, reduction = "umap")

sce <- SingleCellExperiment(assays = list(counts=counts_matrix), colData=coldata)
#reduceDim(sce,"umap")<-umap_embeddings


summed <- aggregateAcrossCells(sce, id=colData(sce)[,"sample"])

#Draw the transcriptomic dissimilarity 
bulk_df <- as.data.frame(assay(summed,"counts"))

dist_matrix  <- 1- cor(bulk_df,method="spearman")

dist_matrix_bis <- dist_matrix * 100
clean_dmat <- as.data.frame(rbind(dist_matrix_bis[3,1:2],
                      dist_matrix_bis[6,4:5],
                      dist_matrix_bis[9,7:8]))

rownames(clean_dmat)<-c("OO100","OO77","OO99")
colnames(clean_dmat)<-c("untreated_vs_in_vitro","untreated_vs_in_vivo")
clean_dmat$patient <-rownames(clean_dmat)
clean_dmat_long <- clean_dmat%>%pivot_longer(cols=c(1:2),names_to="treatment",values_to="distance")
clean_dmat_long$patient <- factor(clean_dmat_long$patient , levels = c("OO77","OO99","OO100"))
```

## Figure 2A: Transcriptomic dissimilarity
```{r fig.width = 4.5,   fig.height = 4.5 }
colors = c( "goldenrod3", "darkcyan","indianred1","tomato", "violet")

plot <- ggplot(clean_dmat_long, aes(x=treatment, y=distance, group=patient)) +
  geom_point(aes(colour = treatment),size= 2, position=position_dodge(width=0)) +
  geom_line( aes(colour = patient), linewidth= 2, alpha=0.5, position=position_dodge(width=0)) +
  scale_x_discrete(labels = c("untreated_vs_in_vitro" = " Ex vivo", "untreated_vs_in_vivo" = "In vivo")) +
  scale_colour_manual(values = colors)+
  ylim(0,8)+
  ylab("Transcriptomic dissimilarity") +
  #+geom_text(aes(label = patient), vjust = -0.3, position = position_dodge(width = 0.55),size= 5) +
  theme_classic()+
  theme(legend.position = "none",
        axis.title.x.bottom=element_blank(),
        axis.text.x.bottom=element_text(size= 16),
        axis.title.y.left=element_text(size= 19),
         axis.text.y.left=element_text(size= 16))
plot
ggsave("Figure_2A_Transcriptomic_dissimilarity.pdf",path= plot_path,width= 6.5,height= 6.5)


#Save data for this Figure
data = plot$data
head(data)
write.table(data , file.path(figures_data, "Figure_2A_Transcriptomic_dissimilarity.txt"), sep = "\t", quote = F)
```

# Differential expression analysis *ex vivo* and *in vivo* samples compared to untreated samples
## Function to perform pseudobulk DE analysis
```{r }
######### Pseudobulk using SCE and cluster of 3 patients
DE_and_FA <- function(patient){
    patient_name = patient
    seurat = subset(seurat_integrated , subset = patient == patient_name)

    DefaultAssay(seurat) <- 'RNA'
    counts_matrix <- GetAssayData(object = seurat, slot = "counts")
    coldata <- seurat@meta.data
    #umap_embeddings<-Embeddings(object = seurat_integrated, reduction = "umap")

    sce <- SingleCellExperiment(assays = list(counts=counts_matrix), colData=coldata)
    #reduceDim(sce,"umap")<-umap_embeddings

    #Aggregate count 
    summed <- aggregateAcrossCells(sce, id=colData(sce)[,c("seurat_clusters","treatment")])
    meta <- colData(summed)  %>% as.data.frame %>% mutate(seurat_clusters = as.character(seurat_clusters))%>%
    mutate(sample_id = paste(orig.ident, seurat_clusters,sep="_"))

    dds <- DESeqDataSetFromMatrix(countData = counts(summed), colData = meta, design = ~ seurat_clusters + treatment)

    # Transform counts for data visualization
    rld <- rlog(dds, blind=TRUE)

    # Run DESeq2 differential expression analysis
    dds <- DESeq(dds)

    
    rld <- rlog(dds, blind=TRUE)
    rld_mat <- assay(rld)
    rld_cor <- cor(rld_mat)   
    rownames(rld_cor) <- meta$sample_id
    colnames(rld_cor) <- meta$sample_id

    #reorder the sample
    meta_reorder <- meta %>% arrange(treatment)
    rld_cor_reorder <- rld_cor[meta_reorder$sample_id, meta_reorder$sample_id]
    anno = as.data.frame(meta_reorder[,"treatment"])
    colnames(anno) = c("treatment")
    rownames(anno) = rownames(rld_cor_reorder)

    ## Define contrasts, extract results table, and shrink the log2 fold changes
    ##vitro
    contrast_vitro <- c("treatment", "vitro", "no")

    res_table_vitro_unshrunken <- results(dds, contrast=contrast_vitro, alpha = 0.05)

    res_table_vitro <- lfcShrink(dds, contrast=contrast_vitro, res=res_table_vitro_unshrunken,type="normal")

    # Turn the results object into a data frame
    res_vitro_df <- data.frame(res_table_vitro)
      #Save this table
    write.table(res_vitro_df,file.path(txt_files,paste0(patient,"_res_vitro_df.txt" )), sep = "\t", quote = F)

    # Subset the significant results
    padj.cutoff = 0.05
    lfc.cutoff= 1
    sig_vitro_res <- filter(res_vitro_df, padj < padj.cutoff & abs(log2FoldChange) > lfc.cutoff) %>% arrange(desc(log2FoldChange )) %>% tibble::rownames_to_column(var="gene")

    
   ##vivo
    contrast_vivo <- c("treatment", "vivo", "no")

    res_table_vivo_unshrunken <- results(dds, contrast=contrast_vivo, alpha = 0.05)

    res_table_vivo <- lfcShrink(dds, contrast=contrast_vivo, res=res_table_vivo_unshrunken,type="normal")

   # Turn the results object into a data frame
    res_vivo_df <- data.frame(res_table_vivo)
   #Save this table
    write.table(res_vivo_df,file.path(txt_files,paste0(patient,"_res_vivo_df.txt" )), sep = "\t", quote = F)
  # Subset the significant results
   padj.cutoff = 0.05
   lfc.cutoff= 1
   sig_vivo_res <- filter(res_vivo_df, padj < padj.cutoff & abs(log2FoldChange) > lfc.cutoff)  %>% arrange(desc(log2FoldChange ))  %>% tibble::rownames_to_column(var="gene")

   length(intersect(sig_vivo_res$gene, sig_vitro_res$gene))

    ### Extract normalized counts for significant genes only

    only_vivo_genes = setdiff(sig_vivo_res$gene,sig_vitro_res$gene)
    all_sig_genes = c(sig_vitro_res$gene,only_vivo_genes )

    normalized_counts <- counts(dds, normalized = TRUE)
    sig_counts <- normalized_counts[rownames(normalized_counts) %in% all_sig_genes, ]

    ## Set a color-blind friendly palette
   heat_colors <- rev(brewer.pal(11, "PuOr"))

   #set colnames of matrix
    colnames(sig_counts) <- meta$sample_id

    #reorder the sample
    meta_reorder <- meta %>% arrange(treatment)
    sig_counts_reorder <-sig_counts[all_sig_genes , meta_reorder$sample_id]
    #use the same anno as the correlation heatmap

    ## Run pheatmap using the metadata data frame for the annotation
     anno_colors = list(treatment = c("no"="deepskyblue2", 
                                      "vitro" = "tomato", 
                                       "vivo" = "plum"))

  for (ext in c("png", "pdf")) {
  pheatmap(sig_counts_reorder, 
           color = heat_colors,
           cluster_col = FALSE, cluster_rows = FALSE, 
           show_rownames = FALSE, show_colnames = FALSE,
           annotation_col = anno, 
           annotation_colors = anno_colors,
           border_color = NA, 
           fontsize = 14, 
           scale = "row", 
           filename = file.path(plot_path, paste0("Figure_Sup_2B_Heatmap_DE_genes_pseudobulk_", patient, ".", ext))) }


  write.table(sig_counts_reorder, file.path(figures_data, paste0( "Figure_Sup_2B_Heatmap_DE_genes_", patient, ".txt")), sep = "\t", quote = F)


  #Scatter plot
  res_vitro_df$gene <- rownames(res_vitro_df)
  res_vivo_df$gene <- rownames(res_vivo_df)
  both <- dplyr::full_join(res_vitro_df,res_vivo_df,by="gene",suffix=c(".vitro",".vivo"))

  sum(is.na(both$padj.vitro))
  both$padj.vitro[is.na(both$padj.vitro)]=1
  sum(is.na(both$padj.vitro))

  sum(is.na(both$padj.vivo))
  both$padj.vivo[is.na(both$padj.vivo)]=1
  sum(is.na(both$padj.vivo))

  sum(is.na(both$log2FoldChange.vitro))
  both$log2FoldChange.vitro[is.na(both$log2FoldChange.vitro)]=0
  sum(is.na(both$log2FoldChange.vitro))

  sum(is.na(both$log2FoldChange.vivo))
  both$log2FoldChange.vivo[is.na(both$log2FoldChange.vivo)]=0
  sum(is.na(both$log2FoldChange.vivo))

  both<-both%>%
          mutate(vitro_sig=ifelse(padj.vitro <= 0.05 & abs(log2FoldChange.vitro) >= 1 ,"yes","no"), 
                vivo_sig=ifelse(padj.vivo <= 0.05 & abs(log2FoldChange.vivo) >= 1,"yes","no"))%>%
                mutate(sig=case_when(
                vitro_sig=="yes" & vivo_sig=="yes" ~ 'both',
                vitro_sig=='yes' & vivo_sig=='no' ~ 'vitro',
                vitro_sig=='no' & vivo_sig=='yes' ~ 'vivo',
                TRUE~'none'))
  correlation_value <- cor(both$log2FoldChange.vitro, both$log2FoldChange.vivo)

  both %>% ggplot(aes(x=log2FoldChange.vitro,y=log2FoldChange.vivo))+
          geom_point(aes(color=sig),size= 1.5 )+
          scale_colour_manual(labels=c("both"="both","none"="none","vitro"="ex vivo","vivo"="in vivo"),
                        values=c("both"="darkseagreen1","none"="grey","vitro"="coral1","vivo"="plum"),name="DE")+
           scale_x_continuous(limits=c(-5,5))+
           scale_y_continuous(limits=c(-12,10))+
           geom_text(x = 5,y = 9.5,label = paste("r=", round(correlation_value, 4)), hjust = 1,vjust = 0, size= 7.5,  fontface ="plain")+
           theme_classic()+  
           ylab("LogFC In vivo vs Untreated")+
           xlab("LogFC Ex vivo vs Untreated")+
          theme(axis.title.x.bottom=element_text(size=22, color = "black"),
                axis.text.x.bottom=element_text(size=22, color = "black"),
                axis.title.y.left=element_text(size=22, color = "black"),
                axis.text.y.left = element_text(size=22, color = "black"),
                legend.position = "None")
  
  ggsave(paste0("Figure_Sup_2A_pseudobulk_Deseq2_scatter_",patient,".pdf"),path=plot_path,width=5,height=5)

  write.table(both, file.path(figures_data, paste0("Figure_Sup_2A_ScatterPlot_Log2FC_", patient, ".txt")), sep = "\t", quote = F)}

patients = c("OO100","OO99","OO77")
all_results =  purrr::map_dfr(patients, DE_and_FA)

```
## Gene set overrepresentative analysis

```{r }

#GO analysis 

Perform_GO_patient <- function(patient){
 DefaultAssay(seurat_integrated)<-"RNA"
 All_genes<- AnnotationDbi::mapIds(org.Hs.eg.db,keys=rownames(seurat_integrated),keytype='SYMBOL',column=c('ENSEMBL'),muliVals=first)
 All_genes<-All_genes[!(is.na(All_genes))]


 H<-msigdbr(species = 'Homo sapiens',category='H')
 H.ensemble<-H%>%dplyr::select(gs_name,ensembl_gene)

 #Function to perform GO with Hallmark and return name of gene set
 perform_GO <- function(genes, universe_genes) {
  enriched_genes <- AnnotationDbi::mapIds(org.Hs.eg.db, keys = genes, keytype = 'SYMBOL', column = 'ENSEMBL', multiVals = 'first')
  enriched_genes <- enriched_genes[!is.na(enriched_genes)]
  enrich_H <- enricher(enriched_genes, universe = All_genes,TERM2GENE = H.ensemble )
  enrich_H_df <- enrich_H@result %>%  filter(p.adjust <= 0.01) 
  enrich_H_df<-enrich_H_df %>% mutate(Description = gsub("HALLMARK_", "", Description))
  return(enrich_H_df) }

padj.cutoff = 0.05
lfc.cutoff= 0.5  #0.75
#vitro
res_vitro_df = read.delim(file.path(txt_files,paste0(patient,"_res_vitro_df.txt" )))
# Subset the significant results

sig_vitro_res <- filter(res_vitro_df, padj < padj.cutoff & abs(log2FoldChange) > lfc.cutoff)  %>% arrange(desc(log2FoldChange ))  %>% tibble::rownames_to_column(var="gene")


vitro_up_genes = sig_vitro_res %>% filter(log2FoldChange > 0) %>% pull(gene)
GO_vitro_up = perform_GO(vitro_up_genes , universe_genes = All_genes)
if(nrow(GO_vitro_up) > 0) {
  GO_vitro_up$Direction = "Up"} else {GO_vitro_up$Direction <- character() }

vitro_down_genes = sig_vitro_res %>% filter(log2FoldChange < 0) %>% pull(gene)
GO_vitro_down = perform_GO(vitro_down_genes , universe_genes = All_genes)

if(nrow(GO_vitro_down) > 0) {
  GO_vitro_down$Direction = "Down"} else {GO_vitro_down$Direction <- character() }
GO_vitro = rbind(GO_vitro_up, GO_vitro_down)
GO_vitro$Comparison = "Ex"

#vivo
res_vivo_df = read.delim(file.path(txt_files,paste0(patient,"_res_vivo_df.txt" )))
sig_vivo_res <- filter(res_vivo_df, padj < padj.cutoff & abs(log2FoldChange) > lfc.cutoff)  %>% arrange(desc(log2FoldChange ))  %>% tibble::rownames_to_column(var="gene")

vivo_up_genes = sig_vivo_res %>% filter(log2FoldChange > 0) %>% pull(gene)
GO_vivo_up = perform_GO (vivo_up_genes , universe_genes = All_genes)

if(nrow(GO_vivo_up) > 0) {
  GO_vivo_up$Direction = "Up"} else {GO_vivo_up$Direction <- character() }
vivo_down_genes = sig_vivo_res %>% filter(log2FoldChange < 0) %>% pull(gene)

vivo_down_genes = sig_vivo_res %>% filter(log2FoldChange < 0) %>% pull(gene)
GO_vivo_down = perform_GO(vivo_down_genes , universe_genes = All_genes)

if(nrow(GO_vivo_down) > 0) {
  GO_vivo_down$Direction = "Down"} else {GO_vivo_down$Direction <- character() }
GO_vivo = rbind(GO_vivo_up, GO_vivo_down)
GO_vivo$Comparison = "In"

GO_patient = rbind(GO_vitro, GO_vivo)
GO_patient$patient = patient

GO_patient}


patients = c("OO100","OO99","OO77")
GO_results <- purrr::map_dfr(patients, Perform_GO_patient)
GO_results$ID <- sub("HALLMARK_","", GO_results$ID )
GO_results$neg_log_adj_p <- -log10(GO_results$p.adjust)
GO_results$comparision_ID <- paste(GO_results$patient, GO_results$Comparison)

ordered_com = c(  "OO77 Ex", "OO99 Ex","OO100 Ex",  "OO77 In", "OO99 In","OO100 In"  )
GO_ordered_pathway = c(  "TNFA_SIGNALING_VIA_NFKB",  "HYPOXIA"      ,"INTERFERON_ALPHA_RESPONSE" , 
                          "UV_RESPONSE_UP" , 
                          "APOPTOSIS"        ,  "P53_PATHWAY"  ,   "INTERFERON_GAMMA_RESPONSE"  , "ESTROGEN_RESPONSE_EARLY" , "CHOLESTEROL_HOMEOSTASIS",  "APICAL_JUNCTION"   , 
                          "KRAS_SIGNALING_UP"  , "EPITHELIAL_MESENCHYMAL_TRANSITION" ,             
                          "ESTROGEN_RESPONSE_LATE",  "GLYCOLYSIS","XENOBIOTIC_METABOLISM", 
                     "MYOGENESIS",    "COMPLEMENT"  ,  "MTORC1_SIGNALING"  , "MITOTIC_SPINDLE","E2F_TARGETS"  , "G2M_CHECKPOINT"   )

GO_results$ID <- factor(GO_results$ID , levels = rev(GO_ordered_pathway))

GO_results$comparision_ID <- factor(GO_results$comparision_ID , levels = ordered_com)
```

## Figure 2B: Bubble plot of GO term 
```{r fig.width = 12,   fig.height = 8 }
plot  <-  ggplot(GO_results, aes(x = comparision_ID, y = ID, color = Direction , size = neg_log_adj_p)) +
  geom_point(alpha = 1, stroke = 1.3) + 
  scale_color_manual(values = c( "#313695", "#A50026"))+
      labs(size = "-log10(p.adjust)") + 
  theme_bw()+
  theme(axis.title.x.bottom=element_blank(),
        axis.text.x.bottom=element_text(size = 18, color = "black"),
        axis.title.y.left=element_blank(),
        axis.text.y.left = element_text(size = 18, color = "black"),
        legend.position = "top",
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 18), 
        legend.box = "vertical",            # Stack legends vertically
        legend.box.just = "left" ,
        legend.spacing.y = unit(2, "pt")         
  ) +
  guides(
    color = guide_legend(nrow = 1),
    size = guide_legend(nrow = 1)
  )

plot
ggsave(filename=file.path(plot_path,paste("Figure_2B_Bubble_GO_DE_BublePlot.pdf")),width= 11,height= 8)

#Save data
data <- plot$data

write.table(data, file.path(figures_data, "Figure_2B_GO_DE_BublePlot.txt"), sep = "\t", quote = F)

```

# Genes in NFkB pathway that exhibit consistent trend across 3 patients 

```{r }
#Combined vivo res table of 3 patient
OO77 = read.delim(file.path(txt_files,"OO77_res_vivo_df.txt" )) 
OO77$patient = "OO77"
OO77$gene = rownames(OO77)
OO77_sig = OO77 %>% filter(padj< 0.05, abs(log2FoldChange)> 0.5)
OO77_up = OO77_sig %>% filter(log2FoldChange > 0) %>% rownames()
OO77_down = OO77_sig %>% filter(log2FoldChange < 0) %>% rownames()

OO99 = read.delim(file.path(txt_files,"OO99_res_vivo_df.txt" )) 
OO99$patient = "OO99"
OO99$gene = rownames(OO99)
OO99_sig = OO99 %>% filter(padj< 0.05, abs(log2FoldChange)> 0.5)
OO99_up = OO99_sig %>% filter(log2FoldChange > 0) %>% rownames()
OO99_down = OO99_sig %>% filter(log2FoldChange < 0) %>% rownames()

OO100 = read.delim(file.path(txt_files,"OO100_res_vivo_df.txt" )) 
OO100$patient = "OO100"
OO100$gene = rownames(OO100)
OO100_sig = OO100 %>% filter(padj< 0.05, abs(log2FoldChange)> 0.5)
OO100_up = OO100_sig %>% filter(log2FoldChange > 0) %>% rownames()
OO100_down = OO100_sig %>% filter(log2FoldChange < 0) %>% rownames()

All = rbind(OO77,OO99,OO100)

#Retrieve gene names
Hallmark_df <- read.delim("/group/poetsch_projects/poetsch_sc/Driver_predict/cancer_gene_lists/h.all.v2023.2.Hs.symbols.gmt")
Hallmark_df[,1] <- sub("HALLMARK_","",Hallmark_df[,1] )

pathway = "TNFA_SIGNALING_VIA_NFKB" 
genes_pathway = Hallmark_df[Hallmark_df[1]== pathway,3:ncol(Hallmark_df)] %>% as.vector()  %>% unlist() %>% as.character()
OO77_up_pathways = intersect(OO77_up ,genes_pathway )
OO99_up_pathways = intersect(OO99_up , genes_pathway )
OO100_up_pathways = intersect(OO100_up , genes_pathway )

OO77_down_pathways = intersect(OO77_down ,genes_pathway )
OO99_down_pathways = intersect(OO99_down , genes_pathway )
OO100_down_pathways = intersect(OO100_down , genes_pathway )

up_intersect = intersect(OO77_up_pathways ,intersect(OO99_up_pathways , OO100_up_pathways))
down_intersect = intersect(OO77_down_pathways ,intersect(OO99_down_pathways , OO100_down_pathways))

intersect = union(up_intersect , down_intersect ) 


# Extract gene expression from single cell data
seurat_sub = subset(seurat_integrated, subset = orig.ident %in% c("OO100-no","OO100-vivo","OO99-no","OO99-vivo","OO77-no","OO77-vivo"))
DefaultAssay(seurat_sub) = "RNA"
options(future.globals.maxSize = 6 * 1024^3)
seurat_sub = SCTransform(seurat_sub,vst.flavor='v1',
                          variable.features.n = nrow(seurat_sub))


seurat_sub$patient = factor(seurat_sub$patient, levels = c("OO77", "OO99","OO100"))
seurat_sub@meta.data = seurat_sub@meta.data %>% arrange(patient)

expression_mat = seurat_sub$SCT$scale.data
expression_intersect = expression_mat[intersect,rownames(seurat_sub@meta.data )]

meta = seurat_sub@meta.data
meta = meta %>% mutate(Treatment = case_when(treatment == "no" ~ "Untreated",
                                            treatment == "vitro" ~ "Ex vivo",
                                            treatment == "vivo" ~ 'In vivo'))
meta = meta %>% mutate(sample = paste(patient, treatment))                                           
seurat_sub@meta.data = meta
rownames(seurat_sub@meta.data) = rownames(meta)

anno_cell = seurat_sub@meta.data[,"Treatment",drop= F]

pathways_color = brewer.pal(4, "Set3")
annotation_colors <- list( Treatment = c("Untreated" = "deepskyblue2", "In vivo" = "violet"))


# Arrange gene by gene expression gap In vivo vs Untreated
expression_intersect_bis <- as.data.frame(t(expression_intersect) )
expression_intersect_bis$CB <- rownames(expression_intersect_bis)
meta$CB = rownames(meta)
expression_intersect_bis <- expression_intersect_bis %>% left_join(meta[,c("CB","patient","Treatment")], by = "CB")
expression_intersect_bis <- expression_intersect_bis %>% pivot_longer(cols = 1:15, values_to = "expression", names_to = "gene")
expression_intersect_bis <- expression_intersect_bis %>% group_by(gene, Treatment) %>% 
                             summarize(mean =  mean(expression)) %>%
                             pivot_wider(names_from = "Treatment", values_from = "mean") %>%
                             mutate(gap = Untreated - `In vivo`) %>%
                             arrange(desc(gap))

gene_order = expression_intersect_bis %>% pull(gene)
expression_intersect = expression_intersect[gene_order,]
```

## Figure 2C: Heatmap of single cell gene expression of common genes in NFKb pathway
```{r }
pheatmap::pheatmap(expression_intersect, 
                   color =  colorRampPalette(rev(brewer.pal(11, "RdYlBu")))(100),
                   show_rownames = TRUE, 
                   show_colnames = FALSE,   
                   breaks = seq(-1, 1, length.out = 101),
                   cluster_row = FALSE, 
                   cluster_col = FALSE, 
                   gaps_col = c( 8558 , 17879 ),
                   annotation_col = anno_cell, 
                   annotation_colors = annotation_colors
                   #,filename = file.path(plot_path, "Figure_2C_Heatmap_common_gene_NFkb.png"),width = 7.5, height = 3.7
                   )

write.table(expression_intersect, file.path(figures_data ,"Figure_2C_Heatmap_common_genes_NFKB.txt"), sep = "\'t", quote = F)
```

```{r include = FALSE}
gap_vector <- expression_intersect_bis$gap
names(gap_vector) <- expression_intersect_bis$gene

gap_df <- tibble(gene = names(gap_vector),
                 gap = gap_vector) %>%
          mutate(gene = factor(gene, levels = rev(gene))) 


# Triangle shape by using geom_area
gap_df$id <- seq_along(gap_vector)
triangle_plot <- ggplot(gap_df, aes(x = gap, y = id)) +
  geom_area(fill = "gray") +
  scale_y_reverse(breaks = NULL) +
  theme_void()

ggsave(file.path(plot_path, "Figure_2C_Triangle.pdf"), width = 0.5,  height = 3.7)
```
```{r}
sessionInfo()
```
