---
title: "Stomach cancer paper Figure 1 BCD"
author: "Vi Dang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    css: font.css
---

```{r setup, include = FALSE}
# Set global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r }
# Load libraries
library(Seurat)
library(tidyverse)
library(RCurl)
library(cowplot)
library(patchwork)
library(stringr)
library(pheatmap)
library(ggplot2)
library(RColorBrewer)
library(SingleCellExperiment)
library(scater)
library(ggpubr)
library(ggtern)
library(variancePartition)
library(edgeR)
```

```{r }
# Specify the path
main_path <- '/group/poetsch_projects/poetsch_sc/Stomach_paper_repository'
rds_path  <- file.path(main_path,'RDS')
txt_files <- file.path(main_path,'txt_files')
plot_path <- file.path(main_path,"Figures")
figures_data <- file.path(main_path,"Figures_data")
dir.create(figures_data)
```
# Figure 1B: UMAP representative all cells by sample of origin
```{r }
# Load seurat object containing data from all 9 samples
seurat <- readRDS(file.path(rds_path,'three_patient_integrated_cca_ref.rds'))
seurat
```

```{r }
#Edit the meta data
seurat$Patient <- str_extract(seurat$orig.ident,"^[^[_-]]*")
seurat$Treatment <- str_extract(seurat$orig.ident,"[^-]*$")

meta <- seurat@meta.data
meta <- meta %>%
  mutate(Sample = case_when(
    grepl("-no", orig.ident) ~ sub("-no", " Untreated", orig.ident),
    grepl("-vitro", orig.ident) ~ sub("-vitro", " Ex vivo", orig.ident),
    grepl("-vivo", orig.ident) ~ sub("-vivo", " In vivo", orig.ident),
    TRUE ~ orig.ident
  ), 
  Treatment = case_when(
    Treatment == "no" ~ "Untreated",
    Treatment == "vitro" ~ "Ex vivo",
    Treatment == "vivo" ~ "In vivo"
   ))  
seurat@meta.data <- meta
```

```{r }
# Normalize count data with SCTransform followed by pca and UMAP on the top 30 PCs
DefaultAssay(seurat)<-'RNA'
options(future.globals.maxSize = 8 * 1024^3)  # 8 GB
seurat <- SCTransform(seurat, vst.flavor = "v1",
                                 #method = "glmGamPoi",
                                 verbose = FALSE)
seurat <- RunPCA(seurat, npcs = 30, )
seurat <- RunUMAP(seurat, reduction = "pca", dims = 1:30, verbose = F)
```

```{r fig.width = 10, fig.height = 8}
seurat$Sample = factor(seurat$Sample, levels = c("OO77 Untreated",  "OO77 Ex vivo", "OO77 In vivo",
                                  "OO99 Untreated", "OO99 Ex vivo", "OO99 In vivo",
                                  "OO100 Untreated", "OO100 Ex vivo", "OO100 In vivo" ))
Idents(seurat) <- "Sample"

color_palette <- c(brewer.pal(12, "Set3")[1:8],  brewer.pal(8, "Set2")[4])

plot <- DimPlot(seurat, reduction = "umap",label=TRUE,label.size = 8, cols= color_palette, repel=T, raster=T, pt.size = 1.5)  + theme_void() +
        theme(legend.text = element_text( size = 15))
plot
ggsave(file.path(plot_path, 'Figure_1B_UMAP_before_integration_all_samples.pdf'),width = 10,height = 8)
```

```{r }
#save plot data 
data = plot$data
colnames(data)[3] = "Sample"
head(data)
write.table(data , file.path(figures_data, "Figure_1B_UMAP_before_integration_all_samples.txt"), sep = "\t", quote = F)
```

# Figure Suplementary 2: Heatmap for top 200 most variable genes

```{r }
#Plot heatmap for top 200 most variable genes

expr_mat <- seurat$RNA$counts
metadata <- seurat@meta.data[,c("Patient","Treatment")]
# Normalize counts using voom
dge <- DGEList(counts = expr_mat)
dge <- calcNormFactors(dge)
expr_voomed <- voom(dge)$E  # Log2-normalized expression

# Assuming expr_voomed is your log2-transformed expression data
gene_variances <- apply(expr_voomed, 1, var)  # Variance for each gene (rows = genes, columns = samples)

# Order the genes by variance (highest variance first)
ordered_genes_var <- sort(gene_variances, decreasing = TRUE)

#Extract gene expression  data
top_genes <- names(ordered_genes_var)[1:200]
expression_matrix <- seurat@assays$SCT$data
expression_matrix <- expression_matrix[rownames(expression_matrix)  %in% top_genes,]
expression_matrix <- expression_matrix[match(top_genes ,rownames(expression_matrix)),]


annotation <- data.frame( 
                    "Treatment"  = seurat@meta.data['Treatment'],
                    "Patient" = seurat@meta.data['Patient'])


# Define colors for annotations
annotation_colors <- list(
  Patient = c("OO77" = "goldenrod3", "OO99" =  "darkcyan", "OO100" = "indianred1"),
  Treatment = c("Untreated" = "deepskyblue", "Ex vivo" = "tomato", "In vivo" = "plum"))

```

```{r fig.width = 9,   fig.height = 5 }
pheatmap(expression_matrix,
         color = viridis::plasma(100),  
         breaks = seq(-2, 2, length.out = 101),
         annotation_col = annotation,  
         annotation_colors = annotation_colors, 
         scale = "row",
         show_colnames = FALSE, show_rownames = FALSE,
         cluster_rows = TRUE, cluster_cols = TRUE
         )
```


# Figure 1C Linear models 
## Fit variance partition separately for Ex vivo and In vivo treatment settings
### Ex vivo
```{r }
#Ex vivo
seurat_ex <- subset(seurat, subset = Treatment %in% c("Untreated", "Ex vivo") )
expr_mat_ex <- seurat_ex$RNA$counts
metadata_ex <- seurat_ex@meta.data[,c("Patient","Treatment")]
# Normalize counts using voom
dge_ex <- DGEList(counts = expr_mat_ex)
dge_ex <- calcNormFactors(dge_ex)
expr_voomed_ex <- voom(dge_ex)$E  # Log2-normalized expression

# Assuming expr_voomed is your log2-transformed expression data
gene_variances_ex <- apply(expr_voomed_ex, 1, var)  # Variance for each gene (rows = genes, columns = samples)

# Order the genes by variance (highest variance first)
ordered_genes_var_ex <- sort(gene_variances_ex, decreasing = TRUE)

# Select the top N most variated genes (e.g., top 2000 genes)
top_genes_ex <- names(ordered_genes_var_ex)[1:2000]
# Subset the expression data to include only the top variated genes
expr_voomed_top_ex <- expr_voomed_ex[top_genes_ex, ]


#Formular: both patient and treatment as fixed effect
formula <- ~ as.factor(Treatment) + as.factor( Patient)  +  as.factor (Patient): as.factor(Treatment)


varPart_ex <- fitExtractVarPartModel(expr_voomed_top_ex, formula, metadata_ex)
colnames(varPart_ex) = c( "Treatment","Patient", "Interaction", "Residuals")
varPart_ex$gene = rownames(varPart_ex)
varPart_ex_long = varPart_ex %>% pivot_longer( cols = 1:4,names_to= "Factor", values_to = "Variance" ) %>% select(gene, Factor, Variance)
varPart_ex_long$Variance = varPart_ex_long$Variance * 100
varPart_ex_long$Factor = factor(varPart_ex_long$Factor , levels = c("Patient", "Treatment", "Interaction", "Residuals"))
```

```{r fig.width = 7,   fig.height =  4}
cols = c("mediumturquoise", "hotpink", "gold", "gray")
plot <- varPart_ex_long %>% ggplot(aes( x= Factor, y = Variance, group = Factor, fill = Factor )) +
            geom_boxplot( width = 0.3, outlier.size = 0.5)+
            geom_violin( width = 1.1, 
                        alpha = 0.5) +    
            geom_jitter(size = 0.1, alpha = 0.05, aes(fill = Factor))+
            scale_fill_manual(values = cols)+
            ylab("% Variance explained")+
            theme_classic() +
            theme(legend.position = "None",
                  axis.text.x.bottom = element_text(size = 20, color = "black"),
                  axis.title.x.bottom = element_text(size = 20, color = "black"),
                  axis.text.y.left = element_text(size = 20, color = "black"),
                  axis.title.y.left = element_text(size = 20, color = "black"))
plot
ggsave(file.path(plot_path , "Figure_1C_VariancePartition_Ex_vivo.pdf"), width = 7, height = 3.7)

#Save data for this Figure
data = plot$data
head(data)
write.table(data , file.path(figures_data, "Figure_1C_VariancePartition_Ex_vivo.txt"), sep = "\t", quote = F)
```
```{r }
# Prepare for Ternary plot
# Modify so that treatment + patient + interaction = 1
varPart_ex <- varPart_ex %>% as.data.frame %>%
  mutate(Total = Treatment + Patient + Interaction) %>%
  mutate(Treatment = Treatment / Total,
         Patient = Patient / Total,
         Interaction = Interaction / Total) %>%
  select(-Total,-Residuals) 

#color dot
patients_ex = varPart_ex %>% filter(Patient > 0.5, Treatment < 0.5, Interaction < 0.5) 
treatments_ex = varPart_ex %>% filter(Patient < 0.5, Treatment > 0.5, Interaction < 0.5) 
interactions_ex = varPart_ex %>% filter(Patient < 0.5, Treatment < 0.5, Interaction > 0.5) 

patient_ex_genes = patients_ex %>% pull(gene)
treatment_ex_genes = treatments_ex %>% pull(gene)
interaction_ex_genes = interactions_ex %>% pull(gene)
```

### In vivo

```{r }
seurat_in <- subset(seurat, subset = Treatment %in% c("Untreated", "In vivo") )
expr_mat_in <- seurat_in$RNA$counts
metadata_in <- seurat_in@meta.data[,c("Patient","Treatment")]
# Normalize counts using voom
dge_in <- DGEList(counts = expr_mat_in)
dge_in <- calcNormFactors(dge_in)
expr_voomed_in <- voom(dge_in)$E  # Log2-normalized expression

# Assuming expr_voomed is your log2-transformed expression data
gene_variances_in <- apply(expr_voomed_in, 1, var)  # Variance for each gene (rows = genes, columns = samples)

# Order the genes by variance (highest variance first)
ordered_genes_var_in <- sort(gene_variances_in, decreasing = TRUE)

# Select the top N most variated genes (e.g., top 2000 genes)
top_genes_in <- names(ordered_genes_var_in)[1:2000]
# Subset the expression data to include only the top variated genes
expr_voomed_top_in <- expr_voomed_in[top_genes_in, ]


#Formular: both patient and treatment as fixed effect
formula <- ~ as.factor(Treatment) + as.factor( Patient)  +  as.factor (Patient): as.factor(Treatment)


varPart_in <- fitExtractVarPartModel(expr_voomed_top_in, formula, metadata_in)
colnames(varPart_in) = c( "Treatment","Patient", "Interaction", "Residuals")
varPart_in$gene = rownames(varPart_in)
varPart_in_long = varPart_in %>% pivot_longer( cols = 1:4,names_to= "Factor", values_to = "Variance" ) %>% select(gene, Factor, Variance)
varPart_in_long$Variance = varPart_in_long$Variance * 100
varPart_in_long$Factor = factor(varPart_in_long$Factor , levels = c("Patient", "Treatment", "Interaction", "Residuals"))
```

```{r fig.width = 7,   fig.height = 4}
cols = c("mediumturquoise", "hotpink", "gold", "gray")
plot <- varPart_in_long %>% ggplot(aes( x= Factor, y = Variance, group = Factor, fill = Factor )) +
            geom_boxplot( width = 0.3, outlier.size = 0.5)+
            geom_violin( width = 1.1, 
                        alpha = 0.5) +    
            geom_jitter(size = 0.1, alpha = 0.05, aes(fill = Factor))+
            scale_fill_manual(values = cols)+
            ylab("% Variance explained")+
            theme_classic() +
            theme(legend.position = "None",
                  axis.text.x.bottom = element_text(size = 20, color = "black"),
                  axis.title.x.bottom = element_text(size = 20, color = "black"),
                  axis.text.y.left = element_text(size = 20, color = "black"),
                  axis.title.y.left = element_text(size = 20, color = "black"))

plot
ggsave(file.path(plot_path , "Figure_1D_variancePartition_In_vivo.pdf"), width = 7, height = 3.7)

#Save data for this Figure
data = plot$data
head(data)
write.table(data , file.path(figures_data, "Figure_1D_VariancePartition_In_vivo.txt"), sep = "\t", quote = F)



```{r }
# Modify so that treatment + patient + interaction = 1
varPart_in <- varPart_in %>% as.data.frame %>%
  mutate(Total = Treatment + Patient + Interaction) %>%
  mutate(Treatment = Treatment / Total,
         Patient = Patient / Total,
         Interaction = Interaction / Total) %>%
  select(-Total, -Residuals)  # Drop the total and residuals column


#color dot
patients_in = varPart_in %>% filter(Patient > 0.5, Treatment < 0.5, Interaction < 0.5) 
treatments_in = varPart_in %>% filter(Patient < 0.5, Treatment > 0.5, Interaction < 0.5) 
interactions_in = varPart_in %>% filter(Patient < 0.5, Treatment < 0.5, Interaction > 0.5) 


patient_in_genes = patients_in %>% pull(gene)
treatment_in_genes = treatments_in %>% pull(gene)
interaction_in_genes = interactions_in %>% pull(gene)

intersect_genes = unique(intersect(c(interaction_in_genes, treatment_in_genes), c(interaction_ex_genes, treatment_ex_genes)))
```

## Ternary plot 
### Ex vivo
```{r fig.width = 12,   fig.height = 10 }
#text label  
patient_label_ex = c()
treatment_label_ex = c("FABP5",  "ME1", "PPP1R14B")
interaction_label_ex = c( "APBB2", "ID1" ,"HNRNPAB" )


#plot
plot <- ggtern(data = varPart_ex, aes(x = Treatment, y = Patient, z = Interaction  )) +
  geom_point(size = 1, alpha = 1, color = "gray") +  # Plot points
  geom_point( data = patients_ex , size = 1, alpha = 1, color = "mediumturquoise",)+
  geom_point( data = treatments_ex, size = 1, alpha = 1, color = "hotpink")+
  geom_point( data = interactions_ex, size = 1, alpha = 1, color = "gold")+
  geom_text(
    data = varPart_ex %>% filter(gene %in% patient_label_ex),
    vjust = 1.1, hjust = 0.57, size = 4, aes(label = gene), color = "red", check_overlap = TRUE) + 
  geom_text(
    data = varPart_ex %>% filter(gene %in% treatment_label_ex),
    vjust = 1.5, hjust = 0.5, size = 4, aes(label = gene), color = "hotpink", check_overlap = TRUE) + 
  geom_text(
    data = varPart_ex %>% filter(gene %in% interaction_label_ex),
    vjust = -0.6, hjust = 1, size = 4, aes(label = gene), color = "gold", check_overlap = TRUE) + 
  # Circle intersect genes with in vivo (below)
  geom_text(
    data = varPart_ex %>% filter(gene %in% setdiff(intersect_genes,c("EPB41L2", "POF1B","ICA1","LGR4"))),
    vjust = -1, hjust = 1.1, size = 4, aes(label = gene), color = "red", check_overlap = TRUE) + 
  geom_text(
    data = varPart_ex %>% filter(gene %in% c("EPB41L2", "POF1B" )),
    vjust = 1.9, hjust = 0.5, size = 4, aes(label = gene), color = "red", check_overlap = TRUE)+
  geom_text(
    data = varPart_ex %>% filter(gene %in% c("ICA1","LGR4")),
    vjust = 1.8, hjust = 0.5, size = 4, aes(label = gene), color = "red", check_overlap = TRUE)+
  geom_point(data = varPart_ex %>% filter(gene %in% intersect_genes ),
             size = 3, shape = 21, stroke = 0.5, fill = NA, color = "red") +
  labs(T = "Patient Effect",
       L = "Treatment Effect",
       R = "Interaction Effect",
       color = "Patient Effect") +
  theme_bw()+  
    theme(
  aspect.ratio = 0.8,
  axis.text = element_text(color = "black"),  # general axis text
  tern.axis.text.T = element_text(size = 16, color = "black"),  # Top axis
  tern.axis.text.L = element_text(size = 16, color = "black"),  # Left axis
  tern.axis.text.R = element_text(size = 16, color = "black")   # Right axis
)

  theme(aspect.ratio = 0.8)  # Makes the triangle taller


plot
ggsave(file.path(plot_path , "Figure_1C_TernaryPlot_Ex_vivo.pdf"), width = 9, height = 8.2)


data <- plot$data
write.table(data , file.path(figures_data, "Figure_1C_TernaryPlot_Ex_vivo.txt"), sep = "\t", quote = F)
```

### In vivo
```{r fig.width = 12,   fig.height = 10}

#text label  
patient_label_in = c()
treatment_label_in = c("GDA", "MBNL2",  "AC012501.2")
interaction_label_in = c()


#plot
plot <- ggtern(data = varPart_in, aes(x = Treatment, y = Patient, z = Interaction  )) +
  geom_point(size = 1, alpha = 1, color = "gray") +  # Plot points
  geom_point( data = patients_in , size = 1, alpha = 1, color = "mediumturquoise",)+
  geom_point( data = treatments_in, size = 1, alpha = 1, color = "hotpink")+
  geom_point( data = interactions_in, size = 1, alpha = 1, color = "gold")+
  geom_text(
    data = varPart_in %>% filter(gene %in% treatment_label_in),
    vjust = -1,hjust = 0.57, size = 4, aes(label = gene), color = "hotpink", check_overlap = TRUE) + 
  geom_text(
    data = varPart_in %>% filter(gene %in% interaction_label_in),
    vjust = -1, hjust = 0.6, size = 4, aes(label = gene), color = "gold", check_overlap = TRUE) + 
  # Label and circle genes intersect with ex vivo
   geom_text(
    data = varPart_in %>% filter(gene %in% setdiff(intersect_genes,c("ICA1","TNFRSF12A", "EPB41L2" ))),
    vjust = -0.7, hjust = 1, size = 4, aes(label = gene), color = "red", check_overlap = TRUE) + 
  geom_text(
    data = varPart_in %>% filter(gene == "TNFRSF12A"),
    vjust = 0, hjust = 1.2, size = 4, aes(label = gene), color = "red", check_overlap = FALSE) + 
  geom_text(
    data = varPart_in %>% filter(gene == "EPB41L2"),
    vjust = 1.7, hjust = 0.5, size = 4, aes(label = gene), color = "red", check_overlap = FALSE) + 
  geom_text(
    data = varPart_in %>% filter(gene == "ICA1"),
    vjust = 0.7, hjust = 1.2, size = 4, aes(label = gene), color = "red", check_overlap = TRUE) + 
  geom_point(data = varPart_in %>% filter(gene %in% intersect_genes),
             size = 3, shape = 21, stroke = 0.5, fill = NA, color = "red") +
  labs(T = "Patient Effect",
       L = "Treatment Effect",
       R = "Interaction Effect",
       color = "Patient Effect") +
  theme_bw()+  
  theme(
  aspect.ratio = 0.8,
  axis.text = element_text(color = "black"),  # general axis text
  tern.axis.text.T = element_text(size = 16, color = "black"),  # Top axis
  tern.axis.text.L = element_text(size = 16, color = "black"),  # Left axis
  tern.axis.text.R = element_text(size = 16, color = "black")   # Right axis
)

  theme(aspect.ratio = 0.8)  # Makes the triangle taller


plot
ggsave(file.path(plot_path , "Figure_1D_TernaryPlot_In_vivo.pdf"), width = 9, height = 8.2)


data <- plot$data
write.table(data , file.path(figures_data, "Figure_1D_TernaryPlot_In_vivo.txt"), sep = "\t", quote = F)

```

```{r}
sessionInfo() 
```